---
title: "p8105_hw3_vas2145"
output: github_document
date: "2023-10-14"
---

#Set up
##Loading libraries and setting standards.
```{r}
library(tidyverse)
library(ggplot2)
library(ggridges)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

#Problem 1
##Loading in Dataset
```{r}
library(p8105.datasets)
data("instacart")

instacart =
  instacart |> 
  as_tibble()
```

The `instacart` dataset includes `r ncol(instacart)` variables of `r nrow(instacart)` observations. The datset includes information such as `order_id`, which is the order number assigned to a particular user, identified by `user_id`. It includes timing of the order, including the day of week (`order_dow), the hour of the order (`order_hour_od_day`) and the amount of time since a previous order was placed by the user (`days_since_prior_order`). It also includes the `product_name`, the product `aisle` and the `department` the product came from. In total, there are 134 aisles, with the fresh vegetables and fresh fruits aisles being the mostly highly ordered from.

```{r}
instacart |> 
  count(aisle_id)
```

```{r}
instacart |> 
  count(aisle_id) |> 
  arrange(desc(n))
```

Plot showing the number of items ordered from each aisle:

```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle", y = "Number of Items") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

A table of the three most popular items ordered from "packaged vegetables fruits", "baking ingredients" and "dog food care" aisles. Each of the products from "packaged vegetables fruits" are substantially more popular than the most popular items from "baking ingredients" and "dog food care" aisles.
```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |> 
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```

On average, ustomers generally order "pink lady apples" earlier in the day than "Coffee Ice Cream," based on the table below.
```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```


#Problem 2

##Loading dataset 

```{r}
library(p8105.datasets)
data("brfss_smart2010")
```

Cleaning data
```{r}
brfss_data =
  brfss_smart2010 |> 
  janitor::clean_names() |> 
  rename(state = locationabbr, county = locationdesc) |> 
  filter(topic == "Overall Health",
    response %in% c("Poor", "Fair", "Good", "Very good", "Excellent")) |> 
  mutate(response = factor(
    response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent")))
```

States that observed at 7 or more locations in 2002:

```{r}
brfss_2002 =
  brfss_data |> 
  filter(year == "2002") |> 
  group_by(state) |> 
  summarise(locations = n_distinct(county)) |> 
  filter(locations >= 7)
```
Connecticut, Florida, Massachusetts, North Carolina, New Jersey, and Pennsylvania all observed at 7 or more locations in 2002.


States that observed at 7 or more locations in 2010:
```{r}
brfss_2010 =
  brfss_data |> 
  filter(year == "2010") |> 
  group_by(state) |> 
  summarise(locations = n_distinct(county)) |> 
  filter(locations >= 7)
```
California, Colorado, Florida, Massachusetts, Maryland, North Carolina, Nebraska, New Jersey, New York, Ohio, Pennsylvania, South Carolina, Texas, and Washington all observed at 7 or more locations in 2010.


#Dataset limited to "Excellent" responses and contains, year, state, and a variable that averages the data_value across locations within a state
```{r}
brfss_ex_data = 
  brfss_data |> 
  filter(response == "Excellent") |> 
  group_by(state, year) |> 
  mutate(avg_data_value = mean(data_value, na.rm = TRUE))
```

A “spaghetti” plot of this average value over time within a state:

```{r}
ggplot(brfss_ex_data, aes(x = year, y = avg_data_value, group = state, color = state)) +
  geom_line() +
  labs(title = "Average State Data Value By Year", x = "Year", y = "Average Data Value", color = "State") +
  theme_minimal()

```

The spaghetti plot shows highly variable data values by each state over the years 2002 to 2010, with a general decrease over time.


Two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State:

```{r}
brfss_ny_data =
  brfss_data |> 
  filter(state == "NY", year %in% c(2006, 2010)) |> 
  separate(county, into = c("state", "county"), "-") |> 
  select(year, county, response, data_value)
  
```

```{r}
ggplot(brfss_ny_data, aes(x = response, y = data_value)) +
  geom_boxplot() +
  geom_point(aes(color = county)) +
  facet_grid(. ~ year) +
  labs(title = "New York State County Distribution of Data Values, 2006 to 2010", x = "Response", y = "Data Value", color = "County") +
  theme_minimal()
```


Problem 3
