---
title: "p8105_hw3_vas2145"
output: github_document
date: "2023-10-14"
---

#Set up
##Loading libraries and setting standards.
```{r}
library(tidyverse)
library(ggplot2)
library(ggridges)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

#Problem 1
##Loading in Dataset
```{r}
library(p8105.datasets)
data("instacart")

instacart =
  instacart |> 
  as_tibble()
```

The `instacart` dataset includes `r ncol(instacart)` variables of `r nrow(instacart)` observations. The datset includes information such as `order_id`, which is the order number assigned to a particular user, identified by `user_id`. It includes timing of the order, including the day of week (`order_dow), the hour of the order (`order_hour_od_day`) and the amount of time since a previous order was placed by the user (`days_since_prior_order`). It also includes the `product_name`, the product `aisle` and the `department` the product came from. In total, there are 134 aisles, with the fresh vegetables and fresh fruits aisles being the mostly highly ordered from.

```{r}
instacart |> 
  count(aisle_id)
```

```{r}
instacart |> 
  count(aisle_id) |> 
  arrange(desc(n))
```

Plot showing the number of items ordered from each aisle:

```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle", y = "Number of Items") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

A table of the three most popular items ordered from "packaged vegetables fruits", "baking ingredients" and "dog food care" aisles. Each of the products from "packaged vegetables fruits" are substantially more popular than the most popular items from "baking ingredients" and "dog food care" aisles.
```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |> 
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```

On average, ustomers generally order "pink lady apples" earlier in the day than "Coffee Ice Cream," based on the table below.
```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```


#Problem 2

##Loading dataset 

```{r}
library(p8105.datasets)
data("brfss_smart2010")
```

Cleaning data
```{r}
brfss_data =
  brfss_smart2010 |> 
  janitor::clean_names() |> 
  rename(state = locationabbr, county = locationdesc) |> 
  filter(topic == "Overall Health",
    response %in% c("Poor", "Fair", "Good", "Very good", "Excellent")) |> 
  mutate(response = factor(
    response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent")))
```

States that observed at 7 or more locations in 2002:

```{r}
brfss_2002 =
  brfss_data |> 
  filter(year == "2002") |> 
  group_by(state) |> 
  summarise(locations = n_distinct(county)) |> 
  filter(locations >= 7)
```
Connecticut, Florida, Massachusetts, North Carolina, New Jersey, and Pennsylvania all observed at 7 or more locations in 2002.


States that observed at 7 or more locations in 2010:
```{r}
brfss_2010 =
  brfss_data |> 
  filter(year == "2010") |> 
  group_by(state) |> 
  summarise(locations = n_distinct(county)) |> 
  filter(locations >= 7)
```
California, Colorado, Florida, Massachusetts, Maryland, North Carolina, Nebraska, New Jersey, New York, Ohio, Pennsylvania, South Carolina, Texas, and Washington all observed at 7 or more locations in 2010.


#Dataset limited to "Excellent" responses and contains, year, state, and a variable that averages the data_value across locations within a state
```{r}
brfss_ex_data = 
  brfss_data |> 
  filter(response == "Excellent") |> 
  group_by(state, year) |> 
  mutate(avg_data_value = mean(data_value, na.rm = TRUE))
```

A “spaghetti” plot of this average value over time within a state:

```{r}
ggplot(brfss_ex_data, aes(x = year, y = avg_data_value, group = state, color = state)) +
  geom_line() +
  labs(title = "Average State Data Value By Year", x = "Year", y = "Average Data Value", color = "State") +
  theme_minimal()

```

The spaghetti plot shows highly variable data values by each state over the years 2002 to 2010, with a general decrease over time.


Two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State:

```{r}
brfss_ny_data =
  brfss_data |> 
  filter(state == "NY", year %in% c(2006, 2010)) |> 
  separate(county, into = c("state", "county"), "-") |> 
  select(year, county, response, data_value)
  
```

```{r}
ggplot(brfss_ny_data, aes(x = response, y = data_value)) +
  geom_boxplot() +
  geom_point(aes(color = county)) +
  facet_grid(. ~ year) +
  labs(title = "New York State County Distribution of Data Values, 2006 to 2010", x = "Response", y = "Data Value", color = "County") +
  theme_minimal()
```


Problem 3

#Import and clean datasets

```{r}
demo_data=
  read_csv(file = "data/nhanes_covar.csv", skip = 4) |> 
  janitor::clean_names() |> 
  filter(age >= 21) |> 
  mutate(
    sex = recode(sex, "1" = "male", "2" = "female"),
    education = recode(education, "1" = "less than high school", "2" = "high school equivalent", "3" = "more than high school")) |> 
  mutate(
    sex = factor(sex),
    education = factor(education)) |> 
  drop_na()
```

```{r}
accel_data =
  read_csv(file = "data/nhanes_accel.csv") |> 
  janitor::clean_names()
```

#merge the datasets:
```{r}
merged_data =
  left_join(demo_data, accel_data, by = "seqn")
```

#Reader-friendly table for the number of men and women in each education category:

```{r}
table_education_sex = merged_data |> 
  group_by(sex, education) |> 
  summarise(count = n(), .groups = "drop") |> 
  pivot_wider(names_from = sex, values_from = count)
  knitr::kable(table_education_sex, caption = "Number of Men and Women in Each Education Category")
```

Most respondents had `more than high school` education, with females having a greater amount of `more than high school` education than males.


#create a visualization of the age distributions for men and women in each education category
```{r}
merged_data |> 
  ggplot(aes(x = education, y = age, fill = sex)) +
    geom_point(alpha = .5) +
    labs(title = "Age Distribution by Sex and Level of Education",
         x = "Level of Education",
         y = "Age") +
   theme_minimal()
```

```{r}
merged_data |> 
  ggplot(aes(x = education, y = age, fill = sex)) +
    geom_boxplot() +
    labs(title = "Age Distribution by Sex and Level of Education",
         x = "Level of Education",
         y = "Age") +
   theme_minimal()
```

Generally, females with more than a high school education are, on average, younger than females with a high school equivalent or less than a high school education, whereas males with less than a high school education are, on average, older than those with a high school equivalent or more than a high school education.


#aggregate across minutes to create a total activity variable for each participant:

```{r}
total_activity_data = merged_data |> 
  mutate(
    total_activity = rowSums(across(c(min1:min45))))
```

#Plot of total activities (y-axis) against age (x-axis), comparing men to women with separate panels for each education level and a trend line to illustrate differences

```{r}
ggplot(total_activity_data, aes(x = age, y = total_activity, color = sex)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_wrap(~ education) +
  labs(title = "Total Activity by Sex and Level of Education",
       x = "Age",
       y = "Total Activity")
```

Across all levels of education, total activity levels decrease from age 20 to age 80. For women, activity levels typically increase between ages 40 and 50, while men typically have decreasing levels of activity between this period.


#three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex
```{r}
time_course_data = total_activity_data |> 
  group_by(education, sex) |> 
  summarize(across(starts_with("min"), ~mean(.), .names = "mean_{.col}")) |> 
  pivot_longer(
    cols = starts_with ("mean"),
    names_to = "time",
    values_to = "mean") |> 
  mutate(time = substring(time, 9),
         time = as.numeric(time))
```

```{r}
time_course_data |> 
  ggplot(aes(x = time, y = mean, color = sex)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(. ~ education) +
  labs(title = "24-hour Activity Time Course by Education Level and Sex",
       x = "Time(min)",
       y = "Mean Activity")
```

```{r}
time_course_data |> 
  ggplot(aes(x = time, y = mean, color = sex)) +
  geom_point(alpha = .5, size = .2) +
  geom_smooth() +
  facet_wrap(. ~ education) +
  labs(title = "24-hour Activity by Education Level and Sex",
       x = "Time(min)",
       y = "Mean Activity")
```

For all levels of education, activity dips in the morning, before increasing at midday. Those with less than a high school education have, on average, greater levels of activity at most time points within the day, with those with a high school equivalent education having, on average, the lowest amount of activity level.
